Scalar and Representative Observables, and Their Associated Statistial Model, for Polarimetric SAR Data

### 2013-11-10 Remote_Sensing_Letter/paper.tex
ready for final round of edition

\documentclass[]{tRSL2e}

@@

%opening
\title{
  Scalar and Representative Observables for Polarimetric SAR Data
}

@@

\begin{abstract}
%Problem statement  
This paper proposes a scalar and representative observable for multi-dimensional POLSAR data,
  from which statistically consistent discrimination measures can be derived.
%What actually were done? and As the result, what is learned?
Specifically, the statistical behaviour of the POLSAR covariance matrix determinant is used
  to derive a scalar and generic statistical model for multi-dimensional POLSAR data,
  which is specifically applicable to the two and three dimensional versions of partial and full monostatic polarimetric SAR data.
As the POLSAR covariance matrix determinant generalizes the SAR intensity towards multiple dimensions,
  the proposed model is able to subsume the traditional SAR intensity model under the umbrella of a unified model. 
%What are the larger implications of the findings presented in this paper?
Consequently, the main beneficial implication of the proposed approach is that
  it provides a consistent theory unifying the currently disconnected proposals for SAR and POLSAR discrimination measures
  which simplifies the adaptation of existing SAR data processing techniques for POLSAR data.
%  it enables the adaptation of many existing SAR data processing techniques for POLSAR data,
%  and at the same time, it provides a consistent theory unifying the seemingly disparate discrimination measure proposals for SAR and POLSAR data.
\end{abstract}

@@

%What is the situation, the background context of your research? the BIG problem!
Relentless growth in computing power has allowed once computationally-demanding Synthetic Aperture Radar (SAR)
to become a feasible and preferred technique for earth observation.
Basic SAR has also been extended in a few directions, one of which is polarimetric SAR (POLSAR).
This exploits the natural polarization property of Electro-Magnetic (EM) waves,
  encoded in multiple channels, compared to traditional one-channel SAR.

%Research Motivation: Why it is important to solve this problem?
%The POLSAR data, however, is multidimensional and stochastic.
POLSAR data, like SAR, is stochastic but is also multi-dimensional, making it even harder to interpret.
It is therefore important to establish a simple and intuitive understanding of the data.
 Statistical models are undoubtedly crucial in understanding its stochastic nature.
While several %multidimensional 
models have been proposed for POLSAR data, they
   tend to be complex and unintuitive due to the multidimensional nature of the data.
Practical POLSAR data processing however, makes heavy use of scalar discrimination measures,
  which should be based on statistically consistent models %for scalar and representative observables 
  of the multidimensional data.
It is thus important to establish scalar and representative observables for multi-dimensional POLSAR data.

@@

%How do others solve this problem?
%The complex POLSAR data have been statistically modelled as following the complex Wishart distribution,
%  which apparently is multidimensional, complex and thus not very intuitive.
A few scalar POLSAR observables with accompanying statistical models have been proposal \citep{Conradsen_2003_TGRS_4, Alberga_2008_IJRS_4129, Joughin_1994_TGRS_562, Lee_1994_TGRS_1017, Touzi_1996_TGRS_519, Lopez-Martinez_2003_TGRS_2232, Erten_2012_Sensors_2766},
  but none is able to provide meaningful scalar discrimination measures.
%  but none of them has led to scalar discrimination measures proposal.
As such no observable has been widely accepted as being highly representative of this multi-dimensional data,
  which severely limits their applicability in practical data processing applications.
Alternatively, a few POLSAR discrimination measures have been proposed,
  but all based on likelihood ratio statistics.
This should ideally be based on an exact and consistent distribution
  but so far only asymptotic distributions have been demonstrated.

%Research Objective: Give your precise problem statement  
This article presents a scalar and representative observable, and its associated generic statistical model, to describe multi-dimensional POLSAR data and provide a consistent foundation for the derivation of discrimination measures.
%Now State your approach  
%This study started off with the observation that the determinant of the covariance matrix is widely used in POLSAR discrimination measures.
%However, our survey of existing scalar models for POLSAR indicates that no statistical model for this observable has been published so far.
%Consequently, in this article, the statistical behaviour of the determinant of POLSAR covariance matrix is further investigated.
%Prof Vuns: comment: the article provides the result, not the journey!
POLSAR data have been statistically modelled as following the complex Wishart distribution.
Consequently, the generic statistical model for the covariance matrix determinant is presented as being just a scalar projection of the multidimensional data.
This model is then used to derive several scalar and consistent statistical descriptions suggesting that their associated observables are capable of being used as discrimination measures for POLSAR.
%
%***IVM: the following sentence doesn't make sense, also, I don't think it's good to talk about 'models of a model'.  Maybe 'model variants' instead?
%%%***IVM***The specific two and three dimensional models of these models, which %also include the derived discrimination measures, are originally designed against both partial and full polarimetric SAR data.

%The specific two and three dimensional models are then validated against both partial and full polarimetric SAR data.
We will also show that %It is also shown that
  the specific one dimensional (1-D) versions of the proposed  models 
  match %perfectly with 
  the traditional statistical model used for SAR intensity.
This effectively incorporates SAR theory under the umbrella of the proposed scalar approach for multi dimensional POLSAR.
The different discrimination proposals for SAR and POLSAR are reviewed in light of this and the  proposed in this paper is shown to provide a strong, unifying and consistent foundation. % relating them all together.
The applicability of these theoretical models will be illustrated by experiments where the specific 1, 2 and 3-D versions of the proposed models are validated against practical data. 
%Furthermore, the approach proposed in this paper
%  not only provides a consistent foundation for the different discrimination measures proposal in POLSAR
%  but also leads to the establishment of a new discriminating measures: the determinant-ratio and the change-ratio.

@@

%**IVM save space!
%\subsection{Scalar Observables for POLSAR Data and their Statistical Models}
%\label{sec:lit_models}

Different target decomposition theorems have identified many possible scalar observables for complex POLSAR data.
\citet{Alberga_2008_IJRS_4129} evaluated the performance of different scalar POLSAR observables for classification.
While many were presented, their corresponding statistical models and classifiers were not available.
Furthermore, the paper concluded that it is impossible to identify a single best representation.
Although, to be fair, the observables were identified for describing a decomposed portion of the complex POLSAR data,
  rather than a unified representation.
%
Using a different approach, given that the joint distribution for POLSAR is known to be the multi-variate complex Wishart,
  it is possible to derive the scalar statistical models for some univariate POLSAR observables.
%This becomes an alternative approach used in the study of POLSAR data.
However, this is non-trivial -- so far, only a handful of such models have been proposed, including:
  \begin{enumerate}
  \item cross-pol ratio $r_{HV/HH} = |S_{HV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol ratio $r_{VV/HH} = |S_{VV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol phase difference $\phi_{VV/HH} = arg(S_{VV}S_{HH}^*) $ \citep{Joughin_1994_TGRS_562} \citep{Lee_1994_TGRS_1017},
  \item magnitude $g=|avg(S_{pq}S_{rs}^*)|$ \citep{Lee_1994_TGRS_1017},
  \item normalized magnitude $\xi = \frac{|avg(S_{pq}S_{rs}^*)|}{\sqrt{avg(|S_{pq}|^2) avg(|S_{rs}|^2)}}$ \citep{Lee_1994_TGRS_1017},
  \item intensity ratio $w = avg(|S_{pq}|^2)/avg(|S_{rs}|^2)$ \citep{Lee_1994_TGRS_1017},
  \item and the Stokes parameters $S_i,0 \leq i \leq 3$ \citep{Touzi_1996_TGRS_519}. 
  \end{enumerate}
More recently, statistical models for
  each element of the POLSAR covariance matrix, i.e. $S_{pq}S_{rs}^*$, \citep{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \citep{Erten_2012_Sensors_2766} have been proposed. Although useful, these have not been shown to result in statistically consistent discrimination measures or 
 be representative of the complex POLSAR data.
%While these models undoubtedly help to further our understandings of the POLSAR data,
 % none of the underlying observables have been shown to meet the dual criteria of
 % 1) resulting in statistically consistent discrimination measures and thus 
 % 2) being representative of the complex POLSAR data.

@@

Euclidean or Manhattan distance measures for matrices are not widely used 
for POLSAR due to the multiplicative nature of the noise.
%, defined respectively as:
%\begin{align}
%  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
%  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
%\end{align}
%where $C_{i,j}$ denotes the $(i,j)$ elements of the POLSAR covariance matrix C,
% $||$ denotes absolute values
%and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
Instead, the Wishart distance is probably most common, as part of the well-known Wishart classifier \citep{Lee_1999_TGRS}, defined \citep{Lee_1994_IJRS_2299} as:
$d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})$
where $tr(C)$ denotes the trace of the POLSAR covariance matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \citep{Anfinsen_2007_ESA_POLINSAR},
\vspace{-2mm}
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
\vspace{-2mm}
the Bartlett distance \citep{Kersten_2005_TGRS_519},
     \vspace{-2mm}
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
     \vspace{-2mm}
the Bhattacharyya distance \citep{Lee_2011_IGARSS_3740},
     \vspace{-2mm}
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
     \vspace{-2mm}
and the Wishart Statistical test distance \citep{Cao_2007_TGRS_3454},
     \vspace{-2mm}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}

Closer examination of these reveal that most are related:
The Bhattacharyya and Bartlett distances are easily shown to be related.
At the same time, Barlett can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among  measures may be due to the fact that
  all of their publications refer to, and are based on, the same statistical model in \citep{Conradsen_2003_TGRS_4}.
In \citep{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrixes $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
     \vspace{-2mm}
\begin{equation}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation}

@@

To detect changes, a test statistic is developed for this discrimination measure.
i.e. a distribution is derived for the dissimilarity measure.
However, \citet{Conradsen_2003_TGRS_4} only use an asymptotic distribution.
By contrast, this paper proposes a statistical model for the determinant of the POLSAR covariance matrix $|C_v|$
  which is capable of providing an exact distribution for the test.

@@

In this section, the generic scalar statistical model for POLSAR is presented.
%In this section, after the basic foundations of POLSAR statistical analysis are introduced,
 % the generic scalar statistical model for the multi-dimensional POLSAR will be presented.
%  this section presents the theoretical model for multi-dimensional POLSAR data. 
%
%In this paper, the POLSAR scattering vector is denoted as $s$.
%For partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
%  the vector is two-dimensional ($d=2$) and is normally written as: 
%\begin{equation}
%s_{part}=\begin{bmatrix}
%S_h\\ 
%S_v
%\end{bmatrix}
%\end{equation}
%For full and monostatic POLSAR data,
 % the vector is three-dimensional ($d=3$) and is presented as:
%\begin{equation}
%s_{full}=\begin{bmatrix}
%S_{hh}\\
%\sqrt{2}S_{hv}\\
%S_{vv}
%\end{bmatrix}
%\end{equation}
%
First, let $\Sigma=E [ss^{*T}]$ denote the population expected value of the POLSAR covariance matrix,
  where $s^{*T}$ is the complex conjugate transpose of the POLSAR scattering vector, $s$. 
%Assuming 
If this is jointly circular complex Gaussian with expected covariance matrix $\Sigma$,
  then the PDF of $s$ can be written as
$pdf(s;\Sigma)=\{1/{\pi^d|\Sigma|}\} exp\{-s^{*T}\Sigma^{-1}s\}$
where $||$ denotes the matrix determinant.
%
The sample POLSAR covariance matrix is formed as the mean of Hermitian outer product of independent single-look scattering vectors,
     \vspace{-4mm}
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $L$ is the number of looks
 and $s_i$ denotes the partial or full POLSAR scattering vector respectively.
%
Complex Wishart distribution statistics are normally used for the scaled covariance matrix
$Z=LC_v$, whose PDF is given as:
\begin{equation}
  pdf(Z;d,\Sigma,L)=\frac{|Z|^{L-d}}{|\Sigma^L|\Gamma_d(L)}e^{-tr(\Sigma^{-1}Z)}
\end{equation}
with $\Gamma_d(L) = \pi^{d(d-1)/2} \prod^{d-1}_{i=0}\Gamma(L-i)$
and $d$ the dimension number of the POLSAR covariance matrix.
%
The approach taken in this paper differs by applying a homoskedastic log transformation  on a less-than-well-known relationship.
\citet{Goodman_1963_AMS_178} found
that the ratio between observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\vspace{-2mm}
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

%We will now use 
We use this to develop a generic scalar statistical model. %log-transformed measures of distance.
From Eqn. \ref{eqn:prod_chi_squared_rv} we have:
\vspace{-2mm}
\begin{eqnarray}
  |C_v| &\sim& |\Sigma_v| \cdot \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi (2L-2i) \label{eqn:determinant_distribution} %\\
%  \ln|C_v| &\sim& \ln|\Sigma_v| - d \cdot \ln(2L) + \sum^{d-1}_{i=0} \Lambda(2L-2i)
%\label{eqn:log_determinant_distribution}  
\end{eqnarray}

Over a homogeneous area, $\Sigma_v$, $d$ and $L$ are considered constant.
Thus Eqn. \ref{eqn:determinant_distribution} indicates that a multiplicative speckle noise pattern is present 
  in the original POLSAR domain.
Moreover, since the average and variance of these chi-squared distributions are %and log-chi-squared distribution
constant, i.e. $avg \left[ \chi(2L) \right] = 2L$ and $var \left[ \chi(2L) \right] = 4L$,
  their product and summation also have fixed summary statistics.
Specifically:
\vspace{-2mm}
{
\begin{align*}
  avg \left[ \prod^{d-1}_{i=0} \chi(2L-2i) \right] &= 2^d \cdot \prod^{d-1}_{i=0} (L-i), \\
  var \left[ \prod^{d-1}_{i=0} \chi(2L-2i) \right] &= \prod^{d-1}_{i=0} 4(L-i)(L-i+1) - \prod^{d-1}_{i=0} 4(L-i)^2, %\\
%  avg \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= d \cdot \ln{2} + \sum^{d-1}_{i=0} \psi^0(L-i), \\
%  var \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= \sum^{d-1}_{i=0} \psi^1(L-i)
\end{align*}
}

>>>>

\documentclass[printer]{tRSL2e}

@@

%opening
\title{
  Scalar and Representative Observables, and Their Associated Statistial Model, for Polarimetric SAR Data
}

@@

\begin{abstract}
%Problem statement  
This paper proposes scalar and representative observables for multi-dimensional POLSAR data,
  which includes several statistically consistent discrimination measures. %can be derived.
%What actually were done? and As the result, what is learned?
Specifically, the statistical behaviour of the POLSAR covariance matrix determinant is used
  to derive a scalar and generic statistical model for multi-dimensional POLSAR data,
  which is shown applicable to the two and three dimensional versions of partial and full monostatic polarimetric SAR data.
As the POLSAR covariance matrix determinant generalizes the SAR intensity towards multiple dimensions,
  the proposed model is able to subsume the traditional SAR intensity model under the umbrella of a unified model. 
%What are the larger implications of the findings presented in this paper?
Consequently, the main beneficial implications are that %of the proposed approach is that
  it provides a consistent theory unifying the currently disconnected proposals for SAR and POLSAR discrimination measures,
  and it simplifies the adaptation of existing SAR data processing techniques for POLSAR data.
%  it enables the adaptation of many existing SAR data processing techniques for POLSAR data,
%  and at the same time, it provides a consistent theory unifying the seemingly disparate discrimination measure proposals for SAR and POLSAR data.
\end{abstract}

@@

%What is the situation, the background context of your research? the BIG problem!
Relentless growth in computing power has allowed the once computationally-demanding Synthetic Aperture Radar (SAR)
to become now a feasible and preferred technique for earth observation.
Basic SAR has also been extended in a few directions, one of which is polarimetric SAR (POLSAR).
POLSAR exploits the natural polarization property of Electro-Magnetic (EM) waves,
  encoded in multiple channels, compared to traditional one-channel SAR.

%Research Motivation: Why it is important to solve this problem?
%The POLSAR data, however, is multidimensional and stochastic.
POLSAR data, like SAR, is stochastic.
Moreover it is also multi-dimensional, making it even harder to interpret.
It is therefore important to establish a simple and intuitive understanding of the data.
 Statistical models are undoubtedly crucial in understanding its stochastic nature.
While several %multidimensional 
models have been proposed for POLSAR data, they
   tend to be complex and unintuitive due to the multidimensional nature of the data.
Practical POLSAR data processing however, makes heavy use of scalar discrimination measures,
  which should be based on statistically consistent models %for scalar and representative observables 
  of the multidimensional data.
It is thus important to establish scalar and representative observables for multi-dimensional POLSAR data.

@@

%How do others solve this problem?
%The complex POLSAR data have been statistically modelled as following the complex Wishart distribution,
%  which apparently is multidimensional, complex and thus not very intuitive.
A few scalar POLSAR observables with accompanying statistical models have been proposed \citep{Conradsen_2003_TGRS_4, Alberga_2008_IJRS_4129, Joughin_1994_TGRS_562, Lee_1994_TGRS_1017, Touzi_1996_TGRS_519, Lopez-Martinez_2003_TGRS_2232, Erten_2012_Sensors_2766},
  but none is able to provide meaningful scalar discrimination measures.
%  but none of them has led to scalar discrimination measures proposal.
As such no observable has been widely accepted as being highly representative of this multi-dimensional data,
  which severely limits their applicability in practical data processing applications.
Alternatively, a few POLSAR discrimination measures have been proposed \citep{Lee_1994_IJRS_2299, Anfinsen_2007_ESA_POLINSAR, Kersten_2005_TGRS_519, Lee_2011_IGARSS_3740, Cao_2007_TGRS_3454, Conradsen_2003_TGRS_4},
  but all are based on likelihood ratio statistics.
These should ideally be based on an exact and consistent distribution
  but so far only asymptotic distributions have been demonstrated \citep{Conradsen_2003_TGRS_4}.

%Research Objective: Give your precise problem statement  
This paper presents scalar and representative observables,
  and one associated generic statistical model,
  that describe multi-dimensional POLSAR data and provide a consistent foundation for the derivation of discrimination measures.
%Now State your approach  
%This study started off with the observation that the determinant of the covariance matrix is widely used in POLSAR discrimination measures.
%However, our survey of existing scalar models for POLSAR indicates that no statistical model for this observable has been published so far.
%Consequently, in this article, the statistical behaviour of the determinant of POLSAR covariance matrix is further investigated.
%Prof Vuns: comment: the article provides the result, not the journey!
%Specifically POLSAR data have been statistically modelled as following the complex Wishart distribution.
The development is based on the common assumption that POLSAR data follows the complex Wishart distribution.  
Subsquently, the generic statistical model for the covariance matrix determinant is presented as being just a scalar projection of the multidimensional data.
This model is then used to derive several scalar and consistent statistical distribution suggesting that their associated observables are capable of being used as discrimination measures for POLSAR.
%
%***IVM: the following sentence doesn't make sense, also, I don't think it's good to talk about 'models of a model'.  Maybe 'model variants' instead?
%%%***IVM***The specific two and three dimensional models of these models, which %also include the derived discrimination measures, are originally designed against both partial and full polarimetric SAR data.

%The specific two and three dimensional models are then validated against both partial and full polarimetric SAR data.
This paper also shows that %It is also shown that
  the specific one dimensional (1-D) version of the proposed  model 
  matches %perfectly with 
  the traditional statistical model used for SAR intensity.
This effectively incorporates the existing SAR theory under the umbrella of the proposed scalar approach for multi dimensional POLSAR.
The different discrimination proposals for SAR and POLSAR are reviewed in light of this and the  proposal in this paper is shown to provide a strong, unifying and consistent foundation. % relating them all together.
The applicability of these theoretical models will be illustrated by experiments where the specific 1, 2 and 3-D versions of the proposed models are validated against real-life captured data. 
%Furthermore, the approach proposed in this paper
%  not only provides a consistent foundation for the different discrimination measures proposal in POLSAR
%  but also leads to the establishment of a new discriminating measures: the determinant-ratio and the change-ratio.

@@

%**IVM save space!
\subsection{Scalar Observables for POLSAR Data} % and their Statistical Models}
\label{sec:lit_models}

Different target decomposition theorems have identified many possible scalar observables for complex POLSAR data.
\citet{Alberga_2008_IJRS_4129} evaluated the performance of different scalar POLSAR observables for classification.
While many observables were presented, their corresponding statistical models and classifiers were not available.
Furthermore, the paper concluded that it is impossible to identify a single best representation,
  although, to be fair, the observables were identified for describing a decomposed portion of the complex POLSAR data,
  rather than a unified representation.

Using a different approach, given that the joint distribution for POLSAR is known to be the multi-variate complex Wishart,
  it is possible to derive the scalar statistical models for some univariate POLSAR observables.
%This becomes an alternative approach used in the study of POLSAR data.
However, this is non-trivial -- and so far, only a handful of such models have been proposed, including:
  \begin{enumerate}
  \item cross-pol ratio $r_{HV/HH} = |S_{HV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol ratio $r_{VV/HH} = |S_{VV}|^2/|S_{HH}|^2$ \citep{Joughin_1994_TGRS_562},
  \item co-pol phase difference $\phi_{VV/HH} = arg(S_{VV}S_{HH}^*) $ \citep{Joughin_1994_TGRS_562} \citep{Lee_1994_TGRS_1017},
  \item magnitude $g=|avg(S_{pq}S_{rs}^*)|$ \citep{Lee_1994_TGRS_1017},
  \item normalized magnitude $\xi = \frac{|avg(S_{pq}S_{rs}^*)|}{\sqrt{avg(|S_{pq}|^2) avg(|S_{rs}|^2)}}$ \citep{Lee_1994_TGRS_1017},
  \item intensity ratio $w = avg(|S_{pq}|^2)/avg(|S_{rs}|^2)$ \citep{Lee_1994_TGRS_1017},
  \item and the Stokes parameters $S_i,0 \leq i \leq 3$ \citep{Touzi_1996_TGRS_519}. 
  \end{enumerate}
More recently, statistical models for
  each element of the POLSAR covariance matrix, i.e. $S_{pq}S_{rs}^*$, \citep{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \citep{Erten_2012_Sensors_2766} have been proposed. Although useful, these have not been shown to result in statistically consistent discrimination measures or 
 to be representative of the complex POLSAR data.
%While these models undoubtedly help to further our understandings of the POLSAR data,
 % none of the underlying observables have been shown to meet the dual criteria of
 % 1) resulting in statistically consistent discrimination measures and thus 
 % 2) being representative of the complex POLSAR data.

@@

Euclidean or Manhattan distance measures for matrices are not widely used 
for POLSAR due to the multiplicative nature of the noise.
%, defined respectively as:
%\begin{align}
%  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
%  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
%\end{align}
%where $C_{i,j}$ denotes the $(i,j)$ elements of the POLSAR covariance matrix C,
% $||$ denotes absolute values
%and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
Instead, the Wishart distance is better known, as part of the Wishart classifier \citep{Lee_1999_TGRS}, defined as \citep{Lee_1994_IJRS_2299}:
$d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})$
where $tr(C)$ denotes the trace of the POLSAR covariance matrix C. 
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \citep{Anfinsen_2007_ESA_POLINSAR},
\vspace{-2mm}
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
\vspace{-2mm}
the Bartlett distance \citep{Kersten_2005_TGRS_519},
     \vspace{-2mm}
  \begin{align}
  d(C_x,C_y) &= 2 \ln |(C_x+C_y)/2| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
     \vspace{-2mm}
the Bhattacharyya distance \citep{Lee_2011_IGARSS_3740},
     \vspace{-2mm}
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
     \vspace{-2mm}
and the Wishart Statistical test distance \citep{Cao_2007_TGRS_3454},
     \vspace{-2mm}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|(L_xC_x+L_yC_y)/(L_x+L_y)| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}

Closer examinations reveal that most are related:
The most obvious are the Bhattacharyya and Bartlett distances. % are easily shown to be related.
At the same time, the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among  measures may be due to the fact that
  all of their publications are based on the same statistical model in \citet{Conradsen_2003_TGRS_4}.
In \citet{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrixes $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
     \vspace{-2mm}
\begin{equation}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation}

@@

To detect changes, a test statistic is developed for this discrimination measure.
i.e. a distribution is derived for the dissimilarity measure.
However, \citet{Conradsen_2003_TGRS_4} only used an asymptotic distribution.
By contrast, this paper proposes a statistical model for the determinant of the POLSAR covariance matrix $|C_v|$
  which is capable of providing an exact distribution for the test.

@@

In this section, the generic scalar statistical model for POLSAR is presented.
%In this section, after the basic foundations of POLSAR statistical analysis are introduced,
 % the generic scalar statistical model for the multi-dimensional POLSAR will be presented.
%  this section presents the theoretical model for multi-dimensional POLSAR data. 
%
%In this paper, the POLSAR scattering vector is denoted as $s$.
%For partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
%  the vector is two-dimensional ($d=2$) and is normally written as: 
%\begin{equation}
%s_{part}=\begin{bmatrix}
%S_h\\ 
%S_v
%\end{bmatrix}
%\end{equation}
%For full and monostatic POLSAR data,
 % the vector is three-dimensional ($d=3$) and is presented as:
%\begin{equation}
%s_{full}=\begin{bmatrix}
%S_{hh}\\
%\sqrt{2}S_{hv}\\
%S_{vv}
%\end{bmatrix}
%\end{equation}
%
First, let $\Sigma=E [ss^{*T}]$ denote the population expected value of the POLSAR covariance matrix,
  where $s^{*T}$ is the complex conjugate transpose of the POLSAR scattering vector, $s$. 
%Assuming 
If this is jointly circular complex Gaussian with expected covariance matrix $\Sigma$,
  then the PDF of $s$ can be written as
$pdf(s;\Sigma)=\{1/{\pi^d|\Sigma|}\} exp\{-s^{*T}\Sigma^{-1}s\}$
where $||$ denotes the matrix determinant.
%
The sample POLSAR covariance matrix is formed as the mean of Hermitian outer product of independent single-look scattering vectors,
     \vspace{-4mm}
\begin{equation}
  C_v = \langle ss^{*T} \rangle = \frac{1}{L} \sum^L_{i=1}s_is_i^{*T}
\end{equation}
where $L$ is the number of looks
 and $s_i$ denotes the partial or full POLSAR scattering vector respectively.
%
Complex Wishart distribution statistics are normally used for the scaled covariance matrix
$Z=LC_v$, whose PDF is given as:
\begin{equation}
  pdf(Z;d,\Sigma,L)=\frac{|Z|^{L-d}}{|\Sigma^L|\Gamma_d(L)}e^{-tr(\Sigma^{-1}Z)}
\end{equation}
with $\Gamma_d(L) = \pi^{d(d-1)/2} \prod^{d-1}_{i=0}\Gamma(L-i)$
and $d$ the dimension number of the POLSAR covariance matrix.
%
The approach taken in this paper differs by making use of applying a less-than-well-known relationship.
\citet{Goodman_1963_AMS_178} found
that the ratio between observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\vspace{-2mm}
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

%We will now use 
This is used to develop a generic scalar statistical model. %log-transformed measures of distance.
From Eqn. \ref{eqn:prod_chi_squared_rv} we have:
\vspace{-2mm}
\begin{eqnarray}
  |C_v| &\sim& |\Sigma_v| \cdot \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi (2L-2i) \label{eqn:determinant_distribution} %\\
%  \ln|C_v| &\sim& \ln|\Sigma_v| - d \cdot \ln(2L) + \sum^{d-1}_{i=0} \Lambda(2L-2i)
%\label{eqn:log_determinant_distribution}  
\end{eqnarray}
